{\bf Demystifying Debugging}

### 2012-11-08 gpi-debug/debugging.tex
minor revise outline
finish debugging slides

	\textbf{``Tell Me a Story''}
	\begin{itemize}
		\item The good debugger's attitude
			% "get more operators"
		\item Binary Search
	\end{itemize}
	\linegap

@@

	{\bf Assorted Tricks}
	\begin{itemize}
		\item Question assumptions
		\item Collect data
	\end{itemize}
	\column{0.45\textwidth}
	%\begin{center}
	%	\includegraphics[width=0.9\textwidth]{vip1066720.jpg}
	%\end{center}
	\end{columns}
\end{frame}

@@

	\linegap
		\texttt{~~~~insert\_sorted(list,~element)~\{} \\
			\texttt{~~~~~~~~\hilight{olivegreen}{int}~i~=~\hilight{brickred}{0};} \\
		\texttt{~~~~~~~~\hilight{brown}{while}~(i~<~length(list))~\{~\hilight{darkcyan}{//~Should~be~"<="!}} \\
		\texttt{~~~~~~~~~~~~\hilight{brown}{if}~(element~<~list[i])~\{} \\
		\texttt{~~~~~~~~~~~~~~~~\hilight{brown}{break};} \\
		\texttt{~~~~~~~~~~~~\}} \\
		\texttt{~~~~~~~~\}} \\
		\texttt{~~~~~~~~insert\_at\_index(list,~element);} \\
		\texttt{~~~~\}} \\
\end{frame}

@@

	\linegap
		\texttt{~~~~input~=~\hilight{brickred}{[99,~42,~3,~100,~87,~56,~22]};} \\
		\texttt{~~~~output~=~\hilight{brickred}{[]};} \\
		\texttt{~~~~\hilight{brown}{foreach}~element~\hilight{brown}{in}~input~\{} \\
		\texttt{~~~~~~~~insert\_sorted(output,~element);} \\
		\texttt{~~~~\}} \\
		\texttt{~~~~\hilight{brown}{return}~is\_sorted(output);~\hilight{darkcyan}{//~returns~false~:(}} \\
	\linegap

@@

	\linegap
		\texttt{~~~~\hilight{brown}{foreach}~element~\hilight{brown}{in}~input~\{} \\
		\texttt{~~~~~~~~\hilight{olivegreen}{print(output);}} \\
		\texttt{~~~~~~~~insert\_sorted(output,~element);} \\
		\texttt{~~~~\}} \\
	\pause
	\linegap

	Creating {\em bounds} on when the problem happened
	\begin{itemize}
		\item Adding more prints = making bounds smaller
		\item (Lots and lots of time? Binary search!)
	\end{itemize}
\end{frame}

@@

	\texttt{~~~~\hilight{brown}{if}~(something\_that\_should\_never\_happen)~\{} \\
	\texttt{~~~~~~~~print(\hilight{brickred}{"it~happened!"});} \\
	\texttt{~~~~~~~~crash();} \\
	\texttt{~~~~\}} \\
\end{frame}
\begin{frame}{Asserts: Stealthier, more concealable prints}
	Adding print statements everywhere is very verbose
	\begin{itemize}
		\item Pro: Get to see everything
		\item Con: Have to look at everything
	\end{itemize}
	15-122's C0 ``contracts'' are assertions
	\begin{itemize}
		\item You don't need to use C0 to write assert-ful code
		\item Asserts can go anywhere, not just around functions/loops
	\end{itemize}
	\linegap

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Assorted Tricks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\breakslide{Assorted Tricks \\ {\normalsize (Useful stuff that isn't really about ``space'' or ``time'')}}

% TODO
\begin{frame}{Rubber Duck / Cardboard Dog Debugging}
	\begin{columns}[l]
	\column{0.75\textwidth}
	Hacker folklore says: describe your problem to a duck!
	\begin{itemize}
		\item ``What could I possibly get out of talking to an inanimate object?''
	\end{itemize}
	\linegap

@@

	Helps figure out what part of the story to start searching at
	}
	\column{0.25\textwidth}
	\includegraphics[width=\textwidth]{rubberduck.jpg}
	\end{columns}
\end{frame}

@@

@@

\begin{frame}{The Toolbox}
	\begin{columns}[l]
	\column{0.05\textwidth}
	\column{0.5\textwidth}
	\textbf{``Tell Me a Story''}
	\begin{itemize}
		\item Get more detail
		\item Rinse \& repeat
		\pause
		\item {\bf Tell a Story of Time}
		\begin{itemize}
			\item Print debugging
			\item Asserts
		\end{itemize}
		\pause
		\item {\bf Tell a Story of Space}
		\begin{itemize}
			\item Minimize the failing test case
			\item Make small tests that work
		\end{itemize}
		\pause
		\item Draw pictures, collect data, question assumptions
	\end{itemize}
	\column{0.45\textwidth}
	\pause
	\begin{center}
		\includegraphics[width=0.9\textwidth]{vip1066720.jpg}
	\end{center}
	\end{columns}
\end{frame}

>>>>

	\textbf{``Tell Me a Story''}
	\begin{itemize}
		\item The good debugger's attitude
			% "get more operators"
	\end{itemize}
	\linegap

@@

	{\bf Assorted Tricks}
	\begin{itemize}
		\item Question assumptions
		\item Draw pictures
		\item Binary search
	\end{itemize}
	\column{0.45\textwidth}
	%\begin{center}
	%	\includegraphics[width=0.9\textwidth]{vip1066720.jpg}
	%\end{center}
	\end{columns}
\end{frame}

@@

	\linegap
		\texttt{~~~~insert\_sorted(list,~element)~\{} \\
			\texttt{~~~~~~~~\hilight{olivegreen}{int}~i~=~\hilight{brickred}{0};} \\
		\texttt{~~~~~~~~\hilight{brown}{while}~(i~<~length(list))~\{~\hilight{darkcyan}{//~Should~be~"<="!}} \\
		\texttt{~~~~~~~~~~~~\hilight{brown}{if}~(element~<~list[i])~\{} \\
		\texttt{~~~~~~~~~~~~~~~~\hilight{brown}{break};} \\
		\texttt{~~~~~~~~~~~~\}} \\
		\texttt{~~~~~~~~\}} \\
		\texttt{~~~~~~~~insert\_at\_index(list,~element,~i);} \\
		\texttt{~~~~\}} \\
\end{frame}

@@

	\linegap
		\texttt{~~~~input~=~\hilight{brickred}{[99,~42,~3,~100,~87,~56,~22,~38]};} \\
		\texttt{~~~~output~=~\hilight{brickred}{[]};} \\
		\texttt{~~~~\hilight{brown}{foreach}~element~\hilight{brown}{in}~input~\{} \\
		\texttt{~~~~~~~~insert\_sorted(output,~element);} \\
		\texttt{~~~~\}} \\
		\texttt{~~~~\hilight{brown}{return}~is\_sorted(output);~\hilight{darkcyan}{//~returns~false~:(}} \\
	\linegap

@@

	\linegap
		\texttt{~~~~\hilight{brown}{foreach}~element~\hilight{brown}{in}~input~\{} \\
		\texttt{~~~~~~~~\hilight{olivegreen}{print(output);}} \\
		\texttt{~~~~~~~~insert\_sorted(output,~element);} \\
		\texttt{~~~~\}} \\

\end{frame}

@@

	\texttt{~~~~\hilight{brown}{if}~(something\_that\_should\_never\_happen)~\{} \\
	\texttt{~~~~~~~~print(\hilight{brickred}{"oh~no,~it~happened!"});} \\
	\texttt{~~~~~~~~crash();} \\
	\texttt{~~~~\}} \\
\end{frame}
\begin{frame}{Asserts: Stealthier, more concealable prints}
	Adding print statements everywhere is very verbose
	\begin{itemize}
		\item Pro: Get to see everything
		\item Con: Have to look at everything
	\end{itemize}
	15-122's C0 ``contracts'' are assertions
	\begin{itemize}
		\item You don't need to use C0 to write assert-ful code
		\item Asserts can go anywhere, not just around functions/loops
	\end{itemize}
	\linegap

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Assorted Tips}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\breakslide{Assorted Tips \\ {\normalsize (Useful stuff that isn't specifically about ``space'' or ``time'')}}

\begin{frame}{Rubber Duck / Cardboard Dog Debugging}
	\begin{columns}[l]
	\column{0.75\textwidth}
	Hacker folklore says: describe your problem to a duck!
	\begin{itemize}
		\item ``What could I possibly get out of talking to an inanimate object?''
	\end{itemize}
	\linegap

@@

	Helps figure out what part of the story to start searching at
	}
	\column{0.25\textwidth}
	\includegraphics[width=\textwidth]{rubberduck.jpg}
	\end{columns}
\end{frame}
\begin{frame}{Rubber Duck / Cardboard Dog Debugging}
	\begin{columns}[l]
	\column{0.75\textwidth}
	{\em Question assumptions} also means {\em question explanations}.
	\begin{itemize}
		\item You: ``I think X is wrong''
		\item Duck: ``How can you be sure?''
	\end{itemize}
	\onslide<2->{
	\linegap
	The {\bf Hypothesis/Test} workflow
	\begin{itemize}
		\item Guess what the root cause is
		\item Write a small test that fails {\em iff} that's right.
		\onslide<3->{
		\begin{itemize}
			\item Not ``change code blindly, see if that fixes it''
			\item Until you know what's wrong, write {\em tests}
		\end{itemize}
		}
	\end{itemize}
	}
	\column{0.25\textwidth}
	\vspace{0.035in}
	\includegraphics[width=\textwidth]{rubberduck.jpg}
	\end{columns}
\end{frame}

@@

\begin{frame}{Binary Search}
	Print statements and unit tests create {\em bounds} on when the problem happened
	\begin{itemize}
		\item Iteratively making bounds smaller helps you hone in on the bug.
		\begin{itemize}
			\item Add more print statements
			\item Write unit tests to test many (small!) parts of the code
		\end{itemize}
		\item Hey wait, this sounds like a certain search algorithm...
	\end{itemize}
	\pause
	Binary search: ``$O(\mathsf{log}(n))$'' is not just talking about code!
	\linegap
	\pause

	\includegraphics[width=\textwidth]{binary1.png}
\end{frame}
\begin{frame}{Binary Search}
	Print statements and unit tests create {\em bounds} on when the problem happened
	\begin{itemize}
		\item Iteratively making bounds smaller helps you hone in on the bug.
		\begin{itemize}
			\item Add more print statements
			\item Write unit tests to test many (small!) parts of the code
		\end{itemize}
		\item Hey wait, this sounds like a certain search algorithm...
	\end{itemize}
	Binary search: ``$O(\mathsf{log}(n))$'' is not just talking about code!
	\linegap

	\includegraphics[width=\textwidth]{binary2.png}
\end{frame}
\begin{frame}{Binary Search}
	Print statements and unit tests create {\em bounds} on when the problem happened
	\begin{itemize}
		\item Iteratively making bounds smaller helps you hone in on the bug.
		\begin{itemize}
			\item Add more print statements
			\item Write unit tests to test many (small!) parts of the code
		\end{itemize}
		\item Hey wait, this sounds like a certain search algorithm...
	\end{itemize}
	Binary search: ``$O(\mathsf{log}(n))$'' is not just talking about code!
	\linegap

	\includegraphics[width=\textwidth]{binary3.png}
\end{frame}
\begin{frame}{Binary Search}
	Print statements and unit tests create {\em bounds} on when the problem happened
	\begin{itemize}
		\item Iteratively making bounds smaller helps you hone in on the bug.
		\begin{itemize}
			\item Add more print statements
			\item Write unit tests to test many (small!) parts of the code
		\end{itemize}
		\item Hey wait, this sounds like a certain search algorithm...
	\end{itemize}
	Binary search: ``$O(\mathsf{log}(n))$'' is not just talking about code!
	\linegap

	\includegraphics[width=\textwidth]{binary4.png}
\end{frame}
\begin{frame}{Binary Search}
	This also applies to ``space'', not just to ''time''
	\begin{itemize}
		\item Sometimes the only thing that makes your code crash is a 1000-line test case
		\begin{itemize}
			\item Binary search on ``which parts of the test need to be there''
		\end{itemize}
	\end{itemize}
	\pause
	\linegap
	Still don't believe me? \texttt{git bisect}
	\begin{itemize}
		\item Helps find which {\em git revision} introduced a bug
		\item Start with one ``known good'' revision, one ``known bad'' one
		\item \texttt{bisect} binary-searches on revisions in-between, asking you at each if it's good or bad.
	\end{itemize}
\end{frame}

\begin{frame}{One Last Aside: Repeatability}
	All the tests and bugs here have been {\em deterministic}.
	\begin{itemize}
		\item If you see the failure once, you can keep making it happen until it's fixed.
	\end{itemize}
	\pause
	\linegap

	In the ``real world'', and in 400-level classes, failures are not so easy to reproduce.
	\begin{itemize}
		\item Consider: \texttt{input = [rand(), rand(), rand(), rand()]}
		\item Might have failed last week, but never since
	\end{itemize}
	\pause
	\linegap
	Make sure your test cases are {\bf repeatable}
	\begin{itemize}
		\item Getting help (from a TA, from a company) easier when you say ``Here is a failing test'' than ''I once saw a test fail''!
		\item Nonrepeatable tests are often complex
		\begin{itemize}
			\item Use the rest of these techniques to produce a relevant, minimal failing test case.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{The Toolbox}
	\begin{columns}[l]
	\column{0.05\textwidth}
	\column{0.5\textwidth}
	\textbf{``Tell Me a Story''}
	\begin{itemize}
		\item Get more detail
		\item Rinse \& repeat
		\pause
		\item {\bf Tell a Story of Time}
		\begin{itemize}
			\item Print debugging
			\item Asserts
		\end{itemize}
		\pause
		\item {\bf Tell a Story of Space}
		\begin{itemize}
			\item Minimize the failing test case
			\item Make small tests that work
		\end{itemize}
		\pause
		\item Draw pictures, collect data, question assumptions, binary search
	\end{itemize}
	\column{0.45\textwidth}
	\pause
	\begin{center}
		\includegraphics[width=0.9\textwidth]{vip1066720.jpg}
	\end{center}
	\end{columns}
\end{frame}
